---
title: "Tassajara Creek H&H Calculations"
output:
  github_document: default
  #html_notebook: default
keep_md: true
---

```{r include=FALSE}
library(tidyverse)
library(lubridate)

sysfonts::font_add(family = "franklin-gothic", 
                   regular = "./fonts/FRABK.TTF", 
                   italic = "./fonts/FRABKIT.TTF", 
                   bold = "./fonts/FRADM.TTF",
                   bolditalic = "./fonts/FRADMIT.TTF")
showtext::showtext_auto()
ggplot2::theme_set(theme_classic() + theme(text = element_text(family = 'franklin-gothic')))
knitr::opts_chunk$set(fig.width=9, fig.height=6) 

``` 

# Flow Hydrographs 

Import Winter 2022-23 storm flow data from [Tassajara Creek below I-580](https://cloud.xylem.com/hydrosphere/public-sites/OWA_1245EDE7887A4C7D888D3671A060A8E1?tab=0&customerId=OWA_1245EDE7887A4C7D888D3671A060A8E1&siteId=TC_BI580):

```{r message=FALSE, warning=FALSE}

import_streamflow_data <- function(csv) {
  return(read_csv(csv) %>%
    janitor::clean_names() %>%
    mutate(result_datetime = lubridate::mdy(result_date) + lubridate::hms(result_time),
           stage_ft = pmax(gh_probe_ft, 0),
           streamflow_cfs = case_when(streamflow_cfs >=0 ~ streamflow_cfs),
           rain_intensity_in = rain_interval_in * 4
           )
    )
}

new_years_storm <- import_streamflow_data("data/tc-bi580_2022-12-31.csv") 

new_years_storm %>% 
  ggplot(aes(y = streamflow_cfs, x = stage_ft)) + 
    geom_line(color = "black") + 
    scale_y_continuous(trans = 'log10', limits = c(10, 1e4)) + 
    scale_x_continuous(trans = 'log10', limits = c(1, 5), breaks = 1:4) +
    ggtitle("Rating Curve for TC_BI580 (Tassajara Creek Below I-580") 
```

Define a function fo fitting power functions:
```{r message=FALSE, warning=FALSE}

power_function_fit <- function(data, y, x) {
  df <- data %>% 
    select({{y}}, {{x}}) %>% 
    filter({{y}} > 0 & {{x}} > 0)
  model <- lm(log(df[[deparse(substitute(y))]]) ~ log(df[[deparse(substitute(x))]]))
  coeff <- summary(model)$coeff
  output <- c(alpha = exp(coeff[1,1]), beta = coeff[2,1])
  power_function <- function(x) {
    return(output["alpha"] * x^output["beta"])
  }
  return(power_function)
}

```

Use the estimated 14 January 2023 peak discharge (via I-580 high water mark) to extrapolate the rating curve:    

```{r message=FALSE, warning=FALSE}

# Original version interpolating using just the stream gauge 
rating_curve <- new_years_storm %>% 
  power_function_fit(y = streamflow_cfs, x = stage_ft)

new_years_storm_pred <- new_years_storm %>%
  #mutate(streamflow_cfs_pred = rating_curve["alpha"] * stage_ft^rating_curve["beta"])
  mutate(streamflow_cfs_pred = rating_curve(stage_ft))

# Updated version of the rating curve using stepwise interpolation based on the I-580 HWM
peak_flow_jan_2023 <- 3524.8

rating_curve <- new_years_storm %>% 
  select(c(streamflow_cfs, stage_ft)) %>% 
  filter(!is.na(streamflow_cfs)) %>% 
  slice_max(stage_ft) %>%
  bind_rows(tribble(
    ~streamflow_cfs, ~stage_ft,
    peak_flow_jan_2023, max(new_years_storm$stage_ft)
  )) %>%
  power_function_fit(y = streamflow_cfs, x = stage_ft)

new_years_storm_pred <- new_years_storm %>%
  mutate(streamflow_cfs_pred = case_when(is.na(streamflow_cfs) ~ rating_curve(stage_ft), TRUE ~ streamflow_cfs),
         is_predicted = is.na(streamflow_cfs))

new_years_storm_pred %>% ggplot(aes(x = stage_ft)) + 
  geom_line(aes(y = streamflow_cfs_pred), color = "red") + 
  geom_line(aes(y = streamflow_cfs), color="black") +
  scale_y_continuous(trans = 'log10', limits = c(10, 1e4)) + 
  scale_x_continuous(trans = 'log10', limits = c(1, 8), breaks = 1:7) +
  ggtitle("Rating curve extrapolated based on January 2023 HWM") + xlab("Stage (ft)") + ylab("Streamflow (cfs)")

ggsave(file="out/gage_rc.svg", plot=last_plot(), width=9, height=6)

```
Plot the winter 2022-23 hydrograph/hyetograph:

```{r message=FALSE, warning=FALSE}
maxRange <- 5000
coeff <- 0.0005

new_years_storm_pred %>% 
  filter(result_datetime < ymd("2023-01-23")) %>% #23")) %>%
  ggplot(aes(x = result_datetime)) +
  geom_area(aes(y = streamflow_cfs_pred, 
                fill = "q_pred")) + 
  geom_area(aes(y = streamflow_cfs, 
                fill = "q")) +
  geom_tile(aes(y = maxRange - rain_intensity_in/coeff/2, 
                height = rain_intensity_in/coeff, 
                color = "precip", fill = "precip")) +
  scale_y_continuous(name = "Streamflow (cfs)",
                     limit = c(0, maxRange),
                     expand = c(0, 0),
                     sec.axis = sec_axis(trans = ~(.-maxRange)*coeff, 
                                         name = "Precipitation (in/hr)")) +
  theme(legend.position = 'none') + 
  #ggtitle("Dec 2022 - Jan 2023 rainfall and streamflow") +
  geom_hline(yintercept = 650 , linetype = "dashed") + annotate("text", x=as.POSIXct(ymd("2023-01-21")), y=100+650 , label="Q2 = 650 cfs") +
  geom_hline(yintercept = 1200, linetype = "dashed") + annotate("text", x=as.POSIXct(ymd("2023-01-21")), y=100+1200, label="Q5 = 1200 cfs") +
  scale_fill_manual(values = c("precip" = "black", "q" = "#499894", "q_pred" = "#86bcb6"),
                    labels = c("precip" = "Precipitation", "q" = "Streamflow (gauged)", "q_pred" = "Streamflow (predicted)"),
                    name = NULL) +
  scale_color_manual(values = c("precip" = "black", "q" = "#499894", "q_pred" = "#86bcb6"),
                    labels = c("precip" = "Precipitation", "q" = "Streamflow (gauged)", "q_pred" = "Streamflow (predicted)"),
                    name = NULL) +
  scale_x_datetime(date_breaks = "days" , date_labels = "%d") + 
  xlab("Date (December 2022 - January 2023)") #\n%b") + 
ggsave(file="out/hyeto_hydro.svg", plot=last_plot(), width=9, height=6)

```
Use the rating curve to estimate the peak flow for 31 December 2022:
```{r message=FALSE, warning=FALSE}
# figure out the peak flow from the December storm
peak_flow_dec_2022 <- new_years_storm_pred %>% 
  filter(date(result_datetime) == ymd("2022-12-31")) %>% 
  pull(streamflow_cfs_pred) %>% 
  max()
print(peak_flow_dec_2022)
```
Convert the 31 December 2022 stage hydrograph to discharge, using the extrapolated rating curve, and modify it to use as upstream boundary condition for HEC-RAS 2D. 
```{r message=FALSE, warning=FALSE}
min_flow <- 100

hydro_dec <- new_years_storm_pred %>% 
  filter(between(result_datetime, ymd("2022-12-31"), ymd("2023-01-02"))) 
ras_hydro_dec <- hydro_dec %>%
  mutate(cfs = case_when(streamflow_cfs_pred > min_flow ~ streamflow_cfs_pred, TRUE ~ min_flow)) %>%
  select(result_datetime, cfs) 

ras_hydro_dec %>% write_csv("data/ras_hydro_2022-12-31.csv")
ggplot() + geom_line(data = ras_hydro_dec, aes(x = result_datetime, y = cfs))
```

Do the same for January (though not actually using the January storm for modeling)
```{r message=FALSE, warning=FALSE}
hydro_jan <- new_years_storm_pred %>% 
  filter(between(result_datetime, ymd("2023-01-14"), ymd("2023-01-18")))
ras_hydro_jan <- hydro_jan %>%
  mutate(cfs = case_when(streamflow_cfs_pred > min_flow ~ streamflow_cfs_pred, TRUE ~ min_flow)) %>%
  select(result_datetime, cfs) 

ras_hydro_jan %>% write_csv("data/ras_hydro_2023-01-14.csv")
ggplot() + geom_line(data = ras_hydro_jan, aes(x = result_datetime, y = cfs)) 
```
Now take the 31 December 2022 hydrograph, and scale it so that the peak is 5200 cfs, the 100-year flow. This will provide an upstream boundary condition for estimating the 100-year flow in HEC-RAS 2D:

```{r message=FALSE, warning=FALSE}
multiplier <- 5200 / max(ras_hydro_dec$cfs)

ras_hydro_q100 <- hydro_dec %>% 
  mutate(cfs = (cfs = case_when(streamflow_cfs_pred * multiplier > min_flow ~ streamflow_cfs_pred * multiplier, TRUE ~ min_flow))) %>%
  select(result_datetime, cfs)

ras_hydro_q100 %>% write_csv("data/ras_hydro_q100.csv")
ggplot() + geom_line(data = ras_hydro_q100, aes(x = result_datetime, y = cfs))

```
Plot all the versions of the hydrograph together:

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = ras_hydro_q100, aes(x = result_datetime, y = cfs ,color = "4. Scaled for Q100(AC)"), linewidth=1) +
  geom_line(data = ras_hydro_dec, aes(x = result_datetime, y = cfs, color = "3. Minimum applied for US BC"), linewidth=1) + 
  geom_line(data = new_years_storm_pred %>% filter(result_datetime >= ymd("2022-12-31") & result_datetime < ymd("2023-01-02")), 
            aes(x = result_datetime, y = streamflow_cfs_pred, color = "2. Extrapolated"), linewidth=1) +
  geom_line(data = new_years_storm_pred %>% filter(result_datetime >= ymd("2022-12-31") & result_datetime < ymd("2023-01-02")), 
            aes(x = result_datetime, y = streamflow_cfs, color = "1. Raw"), linewidth=1) + 
  scale_color_manual(values = c("#499894", "#86bcb6", "black", "#ffae34"), name="") + 
  scale_size_manual(values = c(2, 1, 2, 1)) + 
  #scale_linetype_manual(values = c("solid", "dashed", "dashed", "dashed")) +
  theme(legend.position = "bottom")+
  xlab("") + ylab("Streamflow (cfs)") + scale_x_datetime(position = "top") 
ggsave(file="out/ras_hydro.svg", plot=last_plot(), width=9, height=6)
ggsave(file="out/ras_hydro_short.svg", plot=last_plot(), width=9, height=4) 

```
Examine specific storms and their peaks:
```{r message=FALSE, warning=FALSE}

max_range_left <- 1
max_range_right <- 20
coeff <- max_range_right/max_range_left

new_years_storm_class <- new_years_storm_pred %>%
  filter(result_datetime < ymd("2023-01-23")) %>%
  mutate(storm = as.factor(case_when(
                           result_datetime < ymd_hms("2023-01-02 00:00:00") ~ "1",
                           result_datetime < ymd_hms("2023-01-07 00:00:00") ~ "2",
                           result_datetime < ymd_hms("2023-01-09 18:00:00") ~ NA,
                           result_datetime < ymd_hms("2023-01-13 00:00:00") ~ NA,
                           result_datetime < ymd_hms("2023-01-17 00:00:00") ~ "5",
                             TRUE ~ NA
  )))

new_years_storm_class_summary <- new_years_storm_class %>% group_by(storm) %>% summarize(max_intensity = max(rain_intensity_in),
                                                        total_precip = sum(rain_interval_in),
                                                        max_stage = max(stage_ft),
                                                        max_streamflow = max(streamflow_cfs_pred),
                                                        peak_rainfall_dt = result_datetime[which.max(rain_interval_in)],
                                                        peak_runoff_dt = result_datetime[which.max(stage_ft)],
                                                        #lag_time = as.duration(peak_runoff_dt - peak_rainfall_dt)
                                                        ) %>%
  mutate(storm_label = paste0(
                              total_precip, " in total", "\n", 
                              max_intensity, " in/hr max.", "\n",
                              peak_rainfall_dt
                              ))

new_years_storm_class %>% ggplot(aes(x = result_datetime)) + 
  geom_col(aes(y = rain_intensity_in, color = storm, fill = storm)) +
  geom_line(aes(y = rain_wy_in/coeff)) +
  scale_x_datetime(date_breaks = "days" , date_labels = "%d") + #\n%b") + 
  scale_y_continuous(name = "precipitation intensity (in/hr)",
                   limit = c(0, max_range_left),
                   expand = c(0, 0),
                   sec.axis = sec_axis(trans = ~.*coeff, 
                                       name = "water year cumulative precip (in)")) + 
  geom_point(data = new_years_storm_class_summary %>% filter(storm %in% c(1, 2, 5)), 
             aes(y = max_intensity, x = peak_rainfall_dt, color = storm)) + 
  geom_text(data = new_years_storm_class_summary %>% filter(storm %in% c(1, 2, 5)), 
             aes(y = max_intensity + 0.1, x = peak_rainfall_dt, color = storm, label = storm_label)) +
  theme(legend.position = "none")

#print(sum(new_years_storm_pred$rain_interval_in))
#print(new_years_storm_class_summary)
#print(new_years_storm_pred %>% filter(result_datetime == ymd_hms("2023-01-05 15:30:00")))
#print(new_years_storm_pred %>% filter(result_datetime == ymd_hms("2023-01-06 14:30:00")))
```

# Flow Frequency Analysis

Import peak flows

```{r message=FALSE, warning=FALSE}

# Peak flows estimated via high water marks:
peak_flows <- tribble(
  ~peak_flow_date, ~discharge_cfs,
  ymd("2019-02-13"), 561,
  ymd("2022-11-11"), 105,
  ymd("2022-12-31"), peak_flow_dec_2022,
  ymd("2023-01-14"), peak_flow_jan_2023
)

# Peak flows published for the project with known return intervals:
flow_freq <- tribble(
  ~use, ~name, ~return_interval_y, ~discharge_cfs,
  T, "Q2 BKF",     2,  650, # 18.4 m3/s per MacWilliams et al 2008
  T, "Q25 BKF",   25, 2048, #  58 m3/s per MacWilliams et al 2008
  T, "Q100 BKF", 100, 4300, # 121.8 m3/s per MacWilliams et al 2008
  F, "Q5 Channel Capacity BKF",     5, 1200, #  34 m3/s per MacWilliams et al 2008
  F, "Q100 Alameda County", 100, 5200, # AlCo estimate from Chan and Heard
)
```

Interpolate the Q2, Q25, Q100 peak flow return intervals to estimate a flow frequency discharge relationship: 

```{r message=FALSE, warning=FALSE}

flow_freq_curve <- flow_freq %>% filter(use) %>% 
  power_function_fit(y = discharge_cfs, x = return_interval_y)

flow_freq_curve_inv <- flow_freq %>% filter(use) %>% 
  power_function_fit(y = return_interval_y, x = discharge_cfs)

flow_freq_pred <- flow_freq %>%
  mutate(return_interval_y_pred = flow_freq_curve_inv(discharge_cfs))

peak_flows_pred <- peak_flows %>%
  mutate(return_interval_y_pred = flow_freq_curve_inv(discharge_cfs))

ggplot() + 
  geom_point(data = flow_freq, aes(y = discharge_cfs, x = return_interval_y)) + 
  geom_point(data = peak_flows_pred, aes(y = discharge_cfs, x = return_interval_y_pred), color = "red") + 
  geom_line(data = flow_freq_pred, aes(y = discharge_cfs, x = return_interval_y_pred)) + 
  scale_y_continuous(trans='log10') + 
  scale_x_continuous(trans='log10') + 
  ggtitle("Flow frequency for model profiles") +
  ggrepel::geom_text_repel(data = peak_flows_pred %>% 
                                    rename(return_interval_y = return_interval_y_pred) %>% 
                                    bind_rows(flow_freq) %>% 
                                    mutate(label = case_when(!is.na(peak_flow_date) ~ as.character(peak_flow_date), TRUE ~ name))
                             , aes(y = discharge_cfs, x = return_interval_y, label = label, color = is.na(peak_flow_date)))  +
  scale_color_manual(values=c("TRUE" = "black", "FALSE" = "red"), name="") +
  ylab("Streamflow (cfs)") + xlab("Return Interval (years)") + theme(legend.position = "none")

ggsave(file="out/flowfreq.svg", plot=last_plot(), width=9, height=6) 

```

# Hydraulics

Set up a table of the established cross sections and other notable river stations

```{r message=FALSE, warning=FALSE}
stations <- tribble(
  ~cross_section, ~station_1d, ~station_2d, ~culvert,
   NA,    NA,    0, NA, # US boundary
   NA, 11300,   26, "Gleason",# gleason apron toe
  "B", 10800,  490, NA,    
   NA, 10400,   NA, NA, # pedestrian path
  "D", 10200, 1210, NA,     
  "E",  9300, 2061, NA,     
  "F",  8800, 2568, NA,    
  "G",  8200, 3157, NA,    
  "H",  7600, 3828, NA,  
   NA,  6900,   NA, NA, # 580 frontage apron head
   NA,  6800,   NA, "I-580",# 580 underpass
   NA,    NA, 4507, NA,  # DS boundary
)

# create a table version with just the cross section stations
cross_sections <- stations %>% drop_na(cross_section)

stations
```

Define metric constants to use in sediment transport calculations:

```{r message=FALSE, warning=FALSE}
# gravitational constant, cm/s2
g_cgs <- 981
# grain density and water density, g/cm3
rho_s_cgs <- 2.65
rho_cgs <- 1.00
# kinematic viscosity of water, cm2/s
nu_cgs <- 0.01
```

Define helper functions for the cross section calculations:

```{r message=FALSE, warning=FALSE}

# Takes a raw table of cross section station vs elevation
# Returns a cleaned table of station and elevation with points at regular intervals
prep_xs <- function(data, sta, elev, delta_x) {
  data %>%
    arrange({{sta}}) %>%
    mutate(sta = {{sta}}) %>%
    complete(sta = seq(from = min({{sta}}), to = max({{sta}}), by = delta_x)) %>%
    mutate(gse = zoo::na.approx({{elev}}, x = sta)) %>% 
    select(c(sta, gse))
}

# Takes a cleaned cross section from prep_xs() and a given water surface elevation
# Returns a named vector with the thalweg elevation, depth, cross-sectional area, and wetted perimeter
calc_xs <- function(data, water_elev) {
  data %>% 
    arrange(sta) %>%
    mutate(wse = case_when(water_elev > gse ~ water_elev),
           delta_x = abs(lag(sta, 1) - sta),
           delta_z = abs(lag(gse, 1) - gse),
           depth = wse - gse, 
           hyp_length = sqrt(delta_x^2 + delta_z^2)
           ) %>%
    filter(!is.na(wse)) %>% 
    summarize(thalweg_elevation = min(gse),
              water_surface_elevation = water_elev,
              max_depth = water_surface_elevation - thalweg_elevation,
              cross_sectional_area = sum(delta_x * depth),
              wetted_perimeter = sum(hyp_length),
              ) %>% 
    as.list() %>% 
    list_flatten()
}

# Takes a cleaned cross section from prep_xs() and a given water surface elevation
# Plots the cross section and the water surface at the given elevation
plot_xs <- function(data, water_elev) {
  data %>% 
    arrange(sta) %>%
    mutate(wse = case_when(water_elev > gse ~ water_elev)) %>%
    ggplot(aes(x = sta)) + 
      geom_line(aes(y = gse)) + 
      geom_line(aes(y = wse))
}
  
```


Import surveyed cross section geometries 

```{r message=FALSE, warning=FALSE}
xs_b <- read_csv("data/xs_geom/xs_b.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_d <- read_csv("data/xs_geom/xs_d.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_e <- read_csv("data/xs_geom/xs_e.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_f <- read_csv("data/xs_geom/xs_f.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_g <- read_csv("data/xs_geom/xs_g.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_h <- read_csv("data/xs_geom/xs_h.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

culvert_gleason <- read_csv("data/xs_geom/culvert_gleason.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

culvert_i580 <- read_csv("data/xs_geom/culvert_i580.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

# Combine the cross sections into a single data frame and plot

xs_dfs <- tribble(
  ~xs_id, ~xs_df,
  "B", xs_b,
  "D", xs_d,
  "E", xs_e,
  "F", xs_f,
  "G", xs_g,
  "H", xs_h, 
)

xs_dfs %>% 
  unnest() %>% 
  ggplot(aes(y = gse, x = sta)) + 
    facet_wrap(~xs_id) + 
    geom_line() + 
    ggtitle("Cross section geometries")
```

Example plot with a water surface intersected:

```{r message=FALSE, warning=FALSE}
plot_xs(xs_b, 359)
```

Calculate hydraulic geometries under HWMs (hydraulic radius, etc):

```{r message=FALSE, warning=FALSE}
hwm_geometries <- read_csv("data/high_water_marks_v2.csv") %>%
  janitor::clean_names() %>%
  left_join(xs_dfs, by = join_by(cross_section == xs_id)) %>%
  mutate(result = map2(xs_df, hwm_elevation, calc_xs)) %>% 
  unnest_wider(result)

# The resulting table includes a nested dataframe xs_id with the cross section geometry
hwm_geometries # %>% filter(peak_flow_date == ymd("2022-12-31"))

```

Calculate the minimum channel elevation for each cross section:
```{r message=FALSE, warning=FALSE}
calc_thalweg_elevation <- function(data) {
  min(data$gse)
}

thalweg_elevations <- xs_dfs %>% 
  mutate(thalweg_elevation = map(xs_df, calc_thalweg_elevation)) %>% 
  unnest(thalweg_elevation) %>%
  select(xs_id, thalweg_elevation)

thalweg_elevations
```

Estimate slopes to use in hydraulic calculations (Manning's).

* For channels, use normal depth assumption and calculate average ground slope. 

* For floodplain, use the slopes between the December 2022 HWMs.

```{r message=FALSE, warning=FALSE}
gleason_wse <- seq(from=min(culvert_gleason$gse)+0.1, to=max(culvert_gleason$gse), by=0.1) %>% 
  as_tibble() %>%
  mutate(result = map(value, function(x){calc_xs(data = culvert_gleason, water_elev=x)})) %>% 
  unnest_wider(col = result) %>%
  drop_na() %>%
  mutate(slope = 0.010,
         mannings_n = 0.016,
         hydraulic_radius = cross_sectional_area / wetted_perimeter, 
         discharge_cfs = 1.486 * cross_sectional_area * hydraulic_radius^(2/3) * slope^(1/2) * mannings_n^(-1)) %>%
  bind_rows(tribble(~highwater, ~discharge_cfs, TRUE, peak_flow_dec_2022)) %>%
  arrange(discharge_cfs) %>%
  mutate(culvert_wse = zoo::na.approx(water_surface_elevation)) %>%
  filter(highwater) %>% 
  pull(culvert_wse)

i580_wse <- seq(from=min(culvert_i580$gse)+0.1, to=max(culvert_i580$gse), by=0.1) %>% 
  as_tibble() %>%
  mutate(result = map(value, function(x){calc_xs(data = culvert_i580, water_elev=x)})) %>% 
  unnest_wider(col = result) %>%
  drop_na() %>%
  mutate(slope = 0.010,
         mannings_n = 0.016,
         hydraulic_radius = cross_sectional_area / wetted_perimeter, 
         discharge_cfs = 1.486 * cross_sectional_area * hydraulic_radius^(2/3) * slope^(1/2) * mannings_n^(-1)) %>%
  bind_rows(tribble(~highwater, ~discharge_cfs, TRUE, peak_flow_dec_2022)) %>%
  arrange(discharge_cfs) %>%
  mutate(culvert_wse = zoo::na.approx(water_surface_elevation)) %>%
  filter(highwater) %>% 
  pull(culvert_wse)

slopes_calc <- stations %>%
  filter(!is.na(cross_section) | !is.na(culvert)) %>%
  left_join(hwm_geometries %>% filter(peak_flow_date == ymd("2022-12-31"))) %>%
  select(cross_section, culvert, station_1d, hwm_elevation, thalweg_elevation) %>%
  left_join(tribble(~culvert, ~culvert_wse, ~culvert_gse,
                    "Gleason", gleason_wse, min(culvert_gleason$gse),
                    "I-580", i580_wse, min(culvert_i580$gse),
                    )) %>%
  mutate(hwm_elevation = case_when(is.na(hwm_elevation) ~ culvert_wse, TRUE ~ hwm_elevation),
         thalweg_elevation = case_when(is.na(thalweg_elevation) ~ culvert_gse, TRUE ~ thalweg_elevation)) %>%
  arrange(station_1d) %>%
  mutate(wse_slope_from_ds = (hwm_elevation - lag(hwm_elevation, 1)) / (station_1d - lag(station_1d, 1)),
         gse_slope_from_ds = (thalweg_elevation - lag(thalweg_elevation, 1)) / (station_1d - lag(station_1d, 1))) %>%
  arrange(-station_1d) %>%
  mutate(wse_slope_from_us = (hwm_elevation - lag(hwm_elevation, 1)) / (station_1d - lag(station_1d, 1)),
         gse_slope_from_us = (thalweg_elevation - lag(thalweg_elevation, 1)) / (station_1d - lag(station_1d, 1))) %>%
  mutate(wse_slope_from_ds = case_when(is.na(wse_slope_from_ds) ~ 0.010, TRUE ~ wse_slope_from_ds),
         wse_slope_from_us = case_when(is.na(wse_slope_from_us) ~ 0.010, TRUE ~ wse_slope_from_us),
         gse_slope_from_ds = case_when(is.na(gse_slope_from_ds) ~ 0.010, TRUE ~ gse_slope_from_ds),
         gse_slope_from_us = case_when(is.na(gse_slope_from_us) ~ 0.010, TRUE ~ gse_slope_from_us),
         floodplain = (wse_slope_from_ds + wse_slope_from_us) / 2,
         channel = (gse_slope_from_ds + gse_slope_from_us) / 2
         ) %>%
  # keep just the relevant features
  select(cross_section, channel, floodplain) %>%
  pivot_longer(cols = c(channel, floodplain), names_to = "series", values_to = "slope_calc") %>% 
  drop_na()

slopes_calc
```

Calculate hydraulics based on cross sectional geometries under high water marks and peak flows

```{r message=FALSE, warning=FALSE}
hwm_hydraulics <- hwm_geometries %>% 
  rename(cross_sectional_area_ft2 = cross_sectional_area,
         wetted_perimeter_ft = wetted_perimeter) %>%
  left_join(peak_flows) %>%
  left_join(slopes_calc) %>%
  mutate(slope = slope_calc) %>%
  mutate(velocity_ft_s = discharge_cfs / cross_sectional_area_ft2,
         hydraulic_radius_ft = cross_sectional_area_ft2 / wetted_perimeter_ft,
         mannings_n = 1.486 * cross_sectional_area_ft2 * hydraulic_radius_ft^(2/3) 
                      * slope^(1/2) * discharge_cfs^(-1),
         # metric conversions
         velocity_m_s = velocity_ft_s / 0.3048,
         hydraulic_radius_m = hydraulic_radius_ft / 0.3048,
         cross_sectional_area_m2 = cross_sectional_area_ft2 / 0.3048^2,
         # bed mobilization
         critical_shields_number = 0.15 * slope^(1/4),
         grain_size_mobilized_mm = 10 * rho_cgs * hydraulic_radius_m * slope / 
                         (critical_shields_number * (rho_s_cgs - rho_cgs)),
         grain_size_mobilized_phi = -log2(grain_size_mobilized_mm),
         # suspended transport
         shear_velocity_cm_s = sqrt(g_cgs * (hydraulic_radius_m*100) * slope),
         settling_velocity_ndim = rho_cgs * shear_velocity_cm_s^3 / 
                         ((rho_s_cgs - rho_cgs) * g_cgs * nu_cgs),
         grain_size_suspended_ndim = sqrt(5832 * settling_velocity_ndim),
         grain_size_suspended_mm = 10 * grain_size_suspended_ndim * rho_cgs * nu_cgs^2 /
                         ((rho_s_cgs - rho_cgs) * g_cgs)^(1/3),
         grain_size_suspended_phi = -log2(grain_size_suspended_mm)
  )
hwm_hydraulics

```

```{r message=FALSE, warning=FALSE}
hwm_hydraulics %>% 
  pivot_longer(cols = c(grain_size_suspended_mm, grain_size_mobilized_mm), names_to = "measure", values_to = "grain_size") %>%
  ggplot(aes(x = cross_section, color = series)) + 
  geom_point(aes(y = grain_size, group = series)) +
  scale_color_manual(values = c("channel" = "black", "floodplain" = "red")) + 
  facet_wrap(~measure, ncol = 1) +
  scale_y_continuous(trans = "log2") + 
  ggtitle("Sediment transport estimates based on high water marks")
  
```

# HEC-RAS 1D hydraulic model results

Import and view RAS 1D results

```{r message=FALSE, warning=FALSE}

# RAS 1D result from original geometries
ras_1d <- read_csv("data/hec_ras_1d_out_v3.csv") %>%
  janitor::clean_names() %>% 
  left_join(cross_sections, by = join_by(river_sta == station_1d)) %>%
  mutate(peak_flow_date = dmy(profile)) %>%
  drop_na(reach)

# RAS 1D result from after-storm geometries
ras_1d_afterstorm <- read_csv("data/hec_ras_1d_out_v4_afterstorm.csv") %>%
  janitor::clean_names() %>% 
  left_join(cross_sections, by = join_by(river_sta == station_1d)) %>%
  mutate(peak_flow_date = dmy(profile)) %>%
  drop_na(reach)
# ras_1d_afterstorm_summary <- ras_1d_afterstorm %>% 
#   rename(w_s_elev_ft_afterstorm = w_s_elev_ft, 
#                              min_ch_el_ft_afterstorm = min_ch_el_ft,
#                              ) %>%
#   select(river_sta, profile, w_s_elev_ft_afterstorm, min_ch_el_ft_afterstorm)

# Create a version for just the cross sections
ras_1d_xs <- ras_1d %>% 
  drop_na(cross_section)

# Create a version for just 31 December 2022
hwm_dec_2022 <- hwm_geometries %>% 
  filter(peak_flow_date == "2022-12-31") %>% 
  left_join(cross_sections)

# Summarize the RAS 1D results, reshaping the table of discharge, cross sectional area, and hydraulic radius
ras_1d_pivot <- bind_rows(
  ras_1d_xs %>% pivot_longer(starts_with("q_")) %>% mutate(measure = "discharge_cfs", 
                                                        loc = case_when(name == "q_total_cfs" ~ "total",
                                                                        name == "q_channel_cfs" ~ "channel",
                                                                        name == "q_left_cfs" ~ "left",
                                                                        name == "q_right_cfs" ~ "right"
                                                                        )),
  ras_1d_xs %>% pivot_longer(starts_with("area_")) %>% mutate(measure = "cross_sectional_area_ft2",
                                                           loc = case_when(name == "area_sq_ft" ~ "total",
                                                                           name == "area_channel_sq_ft" ~ "channel",
                                                                           name == "area_left_sq_ft" ~ "left",
                                                                           name == "area_right_sq_ft" ~ "right"
                                                                        )),
  ras_1d_xs %>% pivot_longer(starts_with("hydr_")) %>% mutate(measure = "hydraulic_radius_ft", 
                                                           loc = case_when(name == "hydr_radius_ft" ~ "total",
                                                                           name == "hydr_radius_c_ft" ~ "channel",
                                                                           name == "hydr_radius_l_ft" ~ "left",
                                                                           name == "hydr_radius_r_ft" ~ "right"
                                                                        ))) %>% 
  select(reach, river_sta, profile, measure, loc, value) %>% 
  pivot_wider(names_from = measure, values_from = value) %>%
  left_join(ras_1d %>% select(cross_section, river_sta, profile, min_ch_el_ft, w_s_elev_ft, e_g_slope_ft_ft, q_total_cfs), by = c("river_sta", "profile"))   

ras_1d_pivot
```

Calculate sediment transport estimates for RAS 1D results

````{r message=FALSE, warning=FALSE}
# Use the reshaped table to calculate sediment transport estimates
ras_1d_sed <- ras_1d_pivot %>% 
  drop_na(discharge_cfs) %>%
  mutate(series = case_when((loc == "left" | loc == "right") ~ "floodplain", TRUE ~ loc),
                        velocity_ft_s = discharge_cfs / cross_sectional_area_ft2,
                        slope = e_g_slope_ft_ft, # use this for now
                       # metric conversions
                       velocity_m_s = velocity_ft_s / 0.3048,
                       hydraulic_radius_m = hydraulic_radius_ft / 0.3048,
                       cross_sectional_area_m2 = cross_sectional_area_ft2 / 0.3048^2,
                       # bed mobilization
                       critical_shields_number = 0.15 * slope^(1/4),
                       grain_size_mobilized_mm = 10 * rho_cgs * hydraulic_radius_m * slope / 
                                       (critical_shields_number * (rho_s_cgs - rho_cgs)),
                       grain_size_mobilized_phi = -log2(grain_size_mobilized_mm),
                       # suspended transport
                       shear_velocity_cm_s = sqrt(g_cgs * (hydraulic_radius_m*100) * slope),
                       settling_velocity_ndim = rho_cgs * shear_velocity_cm_s^3 / 
                                       ((rho_s_cgs - rho_cgs) * g_cgs * nu_cgs),
                       grain_size_suspended_ndim = sqrt(5832 * settling_velocity_ndim),
                       grain_size_suspended_mm = 10 * grain_size_suspended_ndim * rho_cgs * nu_cgs^2 /
                                       ((rho_s_cgs - rho_cgs) * g_cgs)^(1/3),
                       grain_size_suspended_phi = -log2(grain_size_suspended_mm)
              )

ras_1d_sed
```

Summarize and plot back-calculated Manning's N estimates

```{r message=FALSE, warning=FALSE}

chan_heard_2006_n = tribble(
  ~cross_section, ~mannings_n,
  "B", 0.0865,
  "D", 0.0368,
  "E", 0.0687,
  "F", 0.0705,
  "G", 0.0520,
  "H", 0.0380
) %>% mutate(series = "floodplain (2006)")

mannings_n <- hwm_hydraulics %>% 
  group_by(series, cross_section) %>%
  summarize(mannings_n = n() / sum(1 / mannings_n)) %>% # harmonic mean
  bind_rows(chan_heard_2006_n)
  
mannings_n %>% 
  ggplot(aes(x = cross_section, y = mannings_n, color = series, label = round(mannings_n,3))) + 
  #geom_point(aes(group = series)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_y_continuous(breaks = c(0, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.09, 0.12), limits = c(0, 0.13), expand=c(0,0)) +
  scale_color_manual(values = c("channel" = "black", "floodplain" = "red", "floodplain (2006)" = "pink"), name="") + 
  ggtitle("Manning's roughness estimates based on high water marks") + 
  geom_hline(yintercept = 0.04, color = "black", linetype = "dashed") +
  geom_hline(yintercept = 0.06, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 0.12, color = "red", linetype = "dashed")   + 
  ylab("Manning's roughness coefficient (n)") + xlab("Cross Section") + 
  theme(legend.position = "top")

ggsave(file="out/mannings.svg", plot=last_plot(), width=9, height=6)

``` 
Calculate and plot bacnkfull discharge

```{r message=FALSE, warning=FALSE}
bankfull <- tribble(
    ~cross_section, ~bankfull_wse, #~w_s_elev_ft, 
    "B", 359.08,
    "D", 356.38,
    "E", 352.73,
    "F", 350.23,
    "G", 345.95,
    "H", 343.16  
    ) %>% 
  left_join(thalweg_elevations, by = join_by(cross_section == xs_id)) %>% 
  mutate(depth = bankfull_wse - thalweg_elevation, bankfull = TRUE) %>%
  rename(min_ch_el_ft = thalweg_elevation, w_s_elev_ft = bankfull_wse)

xs_depth_vs_discharge <- ras_1d %>% drop_na(cross_section) %>%
  left_join(thalweg_elevations, by = join_by(cross_section == xs_id)) %>%
  mutate(depth = w_s_elev_ft - thalweg_elevation) %>%
  select(cross_section, profile, q_total_cfs, depth, min_ch_el_ft, w_s_elev_ft) 

delta_y <- 0.001
elev_levels <- xs_depth_vs_discharge %>% 
  group_by(cross_section) %>% 
  summarize(thalweg_elevation = min(min_ch_el_ft), min_wse = min(w_s_elev_ft), max_wse = max(w_s_elev_ft)) %>%
  mutate(w_s_elev_ft = map2(min_wse, max_wse, function(x, y) seq(from=x+delta_y, to=y-delta_y, by=delta_y))) %>% 
  unnest() %>%
  mutate(depth = w_s_elev_ft - thalweg_elevation)

bankfull_q <-xs_depth_vs_discharge %>%
  bind_rows(elev_levels) %>%
  arrange(cross_section, depth) %>%
  mutate(ln_q_total_cfs = log(q_total_cfs)) %>%
  mutate(ln_q_total_cfs = zoo::na.approx(ln_q_total_cfs)) %>%
  bind_rows(bankfull) %>%
  arrange(cross_section, depth) %>%
  mutate(ln_q_total_cfs = zoo::na.approx(ln_q_total_cfs)) %>%
  mutate(q_total_cfs = exp(ln_q_total_cfs)) %>%
  filter(bankfull) %>%
  mutate(return_interval = flow_freq_curve_inv(q_total_cfs))

ggplot() + 
  geom_point(data = bankfull_q, aes(x = depth, y = q_total_cfs, color = cross_section, shape = "bankfull"), size=2.5) +
  ggrepel::geom_text_repel(data = bankfull_q, aes(x = depth, y = q_total_cfs, color = cross_section, label = round(q_total_cfs))) +
  geom_point(data = xs_depth_vs_discharge, aes(x = depth, y = q_total_cfs, color = cross_section, shape = "modeled"), size=1) + 
  geom_line(data = xs_depth_vs_discharge, aes(x = depth, y = q_total_cfs, color = cross_section), alpha=0.5) + 
  #geom_line(data = xs_depth_vs_discharge, aes(x = depth, y = q_total_cfs, color = cross_section, shape = "modeled")) +
  scale_y_log10() + 
  ylab("Discharge (cfs)") + xlab("Water Depth (ft)") +
  scale_color_manual(values = c("B" = "#ef6f6a", 
                                "D" = "#ffae34",
                                "E" = "#55ad89",
                                "F" = "#c3bc3f",
                                "G" = "#6388b4",
                                "H" = "#8cc2ca"
                                 ), name="Cross Section") +
  scale_shape_manual(values = c("bankfull" = 19, "modeled" = 4), name="")

ggsave(file="out/bankfull.svg", plot=last_plot(), width=9, height=6)
```

Calculate freeboard for each cross section

```{r message=FALSE, warning=FALSE}
capacity <- tribble(
    ~cross_section, ~capacity_wse, 
    "B", 370.05, 					
    "D", 364.88,
    "E", 359.95,
    "F", 356.01,
    "G", 353.96,
    "H", 352.44) 

ras_1d_xs %>%
  left_join(capacity) %>%
  filter(stringr::str_detect(profile, "Q100")) %>% 
  mutate(freeboard = capacity_wse - w_s_elev_ft) %>%
  select(cross_section, profile, freeboard) %>%
  pivot_wider(names_from = profile, values_from = freeboard)
```

Plot velocity and sediment transport estimates based on RAS 1D results

```{r message=FALSE, warning=FALSE}
breaks <- ras_1d_sed %>% group_by(profile, q_total_cfs) %>% summarize() %>% as.list()

ras_1d_sed %>% 
  filter(loc != "total") %>%
  #group_by(series, cross_section, profile, q_total_cfs) %>% 
  #summarize(velocity_ft_s = mean(velocity_ft_s)) %>%
  ggplot(aes(x = q_total_cfs, y = velocity_ft_s, color = cross_section, label = cross_section, shape = loc)) + 
  geom_point() + scale_shape_manual(values=c(19, 24, 25)) +
  facet_grid(cols = vars(series)) + ggtitle("Velocity (via 1D model)") +
  geom_hline(yintercept=10, linetype='dotted', col = 'red') + 
  scale_x_continuous(breaks = breaks$q_total_cfs, labels = breaks$profile, position = "top", name = "", 
                     sec.axis = sec_axis(trans = ~ ., name = "Discharge (cfs)")) + 
  theme(axis.text.x.top = element_text(angle = 45, vjust = 0, hjust=0)) +
  scale_color_manual(values = c("B" = "#ef6f6a",
                                "D" = "#ffae34",
                                "E" = "#55ad89",
                                "F" = "#c3bc3f",
                                "G" = "#6388b4",
                                "H" = "#8cc2ca"
                                 )) + ylab("Velocity (ft/s)")
ggsave(file="out/q_velocity.svg", plot=last_plot(), width=9, height=6)

ras_1d_sed %>% 
  filter(loc != "total") %>%
  #group_by(series, cross_section, profile, q_total_cfs) %>% 
  #summarize(grain_size_mobilized_mm = mean(grain_size_mobilized_mm)) %>%
  ggplot(aes(x = q_total_cfs, y = grain_size_mobilized_mm, color = cross_section, label = cross_section, shape = loc)) + 
  geom_point() + scale_shape_manual(values=c(19, 24, 25)) +
  facet_grid(cols = vars(series)) + ggtitle("Bed Mobilization (via 1D model)") +
  scale_y_continuous(limits = c(0,20), breaks = c(2^0, 2^1, 2^2, 2^3, 2^4)) + 
    scale_x_continuous(breaks = breaks$q_total_cfs, labels = breaks$profile, position = "top", name = "", 
                     sec.axis = sec_axis(trans = ~ ., name = "Discharge (cfs)")) + 
  geom_hline(yintercept=2^0, col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^1,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^2,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^3,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^4 ,  col = "gray", linetype = "dotted") +
  theme(axis.text.x.top = element_text(angle = 45, vjust = 0, hjust=0)) +
  scale_color_manual(values = c("B" = "#ef6f6a",
                                "D" = "#ffae34",
                                "E" = "#55ad89",
                                "F" = "#c3bc3f",
                                "G" = "#6388b4",
                                "H" = "#8cc2ca"
                                 )) + ylab("Grain Size (mm)")

ggsave(file="out/q_sed_bedmobiliz.svg", plot=last_plot(), width=9, height=6)

ras_1d_sed %>% 
  filter(loc != "total") %>%
  #group_by(series, cross_section, profile, q_total_cfs) %>% 
  #summarize(grain_size_suspended_mm = mean(grain_size_suspended_mm)) %>%
  ggplot(aes(x = q_total_cfs, y = grain_size_suspended_mm, color = cross_section, label = cross_section, shape = loc)) + 
  geom_point() + scale_shape_manual(values=c(19, 24, 25)) +
  facet_grid(cols = vars(series)) + ggtitle("Suspended Transport (via 1D model)") +
  scale_y_continuous(limits = c(0,2.25), breaks = c(2^-4, 2^-3, 2^-2, 2^-1, 2^0, 2^1)) +
    scale_x_continuous(breaks = breaks$q_total_cfs, labels = breaks$profile, position = "top", name = "", 
                     sec.axis = sec_axis(trans = ~ ., name = "Discharge (cfs)")) + 
  geom_hline(yintercept=2^-4, col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^-3,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^-2,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^-1,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^0 ,  col = "gray", linetype = "dotted") +
  geom_hline(yintercept=2^1 ,  col = "gray", linetype = "dotted") +
  theme(axis.text.x.top = element_text(angle = 45, vjust = 0, hjust=0)) +
  scale_color_manual(values = c("B" = "#ef6f6a",
                                "D" = "#ffae34",
                                "E" = "#55ad89",
                                "F" = "#c3bc3f",
                                "G" = "#6388b4",
                                "H" = "#8cc2ca"
                                 )) + ylab("Grain Size (mm)")

ggsave(file="out/q_sed_suspended.svg", plot=last_plot(), width=9, height=6)

```
Variation on the velocity plot combined on a single axis

```{r message=FALSE, warning=FALSE}
ras_1d_sed %>% 
  filter(loc != "total") %>%
  #group_by(series, cross_section, profile, q_total_cfs) %>% 
  #summarize(velocity_ft_s = mean(velocity_ft_s)) %>%
  ggplot(aes(x = q_total_cfs, y = velocity_ft_s)) + 
  geom_point(aes(color = cross_section, shape = series)) + scale_shape_manual(values=c(19, 21)) +
  #geom_smooth(aes(color = cross_section, shape = series)) + scale_shape_manual(values=c(19, 21)) +
  #facet_grid(cols = vars(series)) + 
  ggtitle("Velocity (via 1D model)") +
  scale_x_continuous(breaks = breaks$q_total_cfs, labels = breaks$profile, position = "top", name = "", 
                     sec.axis = sec_axis(trans = ~ ., name = "Discharge (cfs)")) + 
  theme(axis.text.x.top = element_text(angle = 45, vjust = 0, hjust=0), panel.grid.major.y = element_blank()) + 
  geom_hline(yintercept=1.5, col = "gray", linetype = "dotted") +
  geom_hline(yintercept=3,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=4,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=6,   col = "gray", linetype = "dotted") +
  geom_hline(yintercept=10,  col = "gray", linetype = "dotted") +
  scale_y_continuous(breaks = c(0, 1.5, 3, 4, 6, 10)) +
  scale_color_manual(values = c("B" = "#ef6f6a",
                                "D" = "#ffae34",
                                "E" = "#55ad89",
                                "F" = "#c3bc3f",
                                "G" = "#6388b4",
                                "H" = "#8cc2ca"
                                 )) + ylab("Velocity (ft/s)")

ggsave(file="out/q_velocity_stacked.svg", plot=last_plot(), width=9, height=6)

```
Longitudinal profile plot of RAS 1D results

```{r message=FALSE, warning=FALSE}
profile_colors <- c("Q100 AC" = "#8cc2ca",
                                "Q100 BKF" = "#8cc2ca",
                                "Q25 BKF" = "#8cc2ca",
                                "Q5 Channel" = "#8cc2ca",
                                "Q2 BKF" = "#8cc2ca",
                                "14 Jan 2023" = "#ffae34",
                                "31 Dec 2022" = "#bb7693",
                                "13 Feb 2019" = "#c3bc3f",
                                "11 Nov 2022" = "#55ad89"
                                 )

ras_1d %>% 
  left_join(hwm_geometries) %>% 
  filter(profile != "5 Jan 1982") %>%
  left_join(bankfull %>% rename(bankfull_wse = w_s_elev_ft) %>% select(cross_section, bankfull_wse)) %>%
  arrange(water_surface_elevation) %>%
  mutate(profile_fct = forcats::fct_reorder(profile, -w_s_elev_ft)) %>%
  ggplot(aes(x = river_sta)) + 
    geom_line(aes(y = w_s_elev_ft, color = profile_fct) ) + 
    geom_line(aes(y = min_ch_el_ft), color = "black") +
    scale_x_reverse() + theme(legend.title=element_blank()) + 
    geom_point(aes(y = water_surface_elevation, color = profile_fct), shape = 9, size = 4) + 
    geom_text(aes(y = bankfull_wse, label = "- -"), color = "black", size = 4) +
    geom_text(data = left_join(hwm_dec_2022, capacity), aes(x = station_1d, y = capacity_wse, label = "—"), color = "black", size = 4) +
  scale_color_manual(values = profile_colors) +
    #ggtitle("2D model result") + 
  xlab("Station (ft) (Zone 7)") + 
  ylab("Elevation (ft NGVD29)") +
  geom_text(data = hwm_dec_2022, aes(x = station_1d, 
                                     y = thalweg_elevation - 1, 
                                     label = cross_section ))
#
ggsave(file="out/lp_1d_wse.svg", plot=last_plot(), width=9, height=6)
```
Now comparing against 2023 results 
```{r message=FALSE, warning=FALSE}
ras_1d %>% 
  bind_rows(ras_1d_afterstorm, .id="geometry") %>%
  mutate(depth = w_s_elev_ft - min_ch_el_ft) %>%
  select(geometry, river_sta, cross_section, profile, w_s_elev_ft, min_ch_el_ft, depth) %>%
  mutate(profile_fct = forcats::fct_reorder(profile, -w_s_elev_ft)) %>%
  pivot_wider(names_from = geometry, values_from = c(min_ch_el_ft, w_s_elev_ft, depth)) %>%
  ggplot(aes(x = river_sta)) + 
  geom_line(aes(y = w_s_elev_ft_1, color = profile_fct)) + 
  geom_line(aes(y = w_s_elev_ft_2, color = profile_fct), linetype="dashed") +
  geom_line(aes(y = min_ch_el_ft_1)) + 
  geom_line(aes(y = min_ch_el_ft_2), linetype="dashed") + scale_color_manual(values = profile_colors) + 
      scale_x_reverse() + theme(legend.title=element_blank()) +
  xlab("Station (ft) (Zone 7)") + 
  ylab("Elevation (ft NGVD29)") + 
  geom_text(data = left_join(hwm_dec_2022, capacity), aes(x = station_1d, y = capacity_wse, label = "—"), color = "black", size = 4) +
  geom_text(data = hwm_dec_2022, aes(x = station_1d, y = thalweg_elevation - 2.5, label = cross_section )) 


ggsave(file="out/lp_1d_wse_afterstorm.svg", plot=last_plot(), width=9, height=6)

```
Now comparing terrain differences only
```{r message=FALSE, warning=FALSE}
ras_1d %>% 
  bind_rows(ras_1d_afterstorm, .id="geometry") %>%
  mutate(depth = w_s_elev_ft - min_ch_el_ft) %>%
  select(geometry, river_sta, cross_section, profile, w_s_elev_ft, min_ch_el_ft, depth) %>%
  mutate(profile_fct = forcats::fct_reorder(profile, -w_s_elev_ft)) %>%
  pivot_wider(names_from = geometry, values_from = c(min_ch_el_ft, w_s_elev_ft, depth)) %>%
  ggplot(aes(x = river_sta)) + 
  geom_line(aes(y = min_ch_el_ft_1)) + 
  geom_line(aes(y = min_ch_el_ft_2), linetype="dashed") + scale_color_manual(values = profile_colors) + 
      scale_x_reverse() + theme(legend.title=element_blank()) +
  xlab("Station (ft) (Zone 7)") + 
  ylab("Elevation (ft NGVD29)") + 
  geom_text(data = hwm_dec_2022, aes(x = station_1d, y = thalweg_elevation - 2.5, label = cross_section )) + 
  geom_line(data = left_join(hwm_dec_2022, capacity), aes(x = station_1d, y = capacity_wse), color = "black", linetype="dotted") +
  geom_line(data = left_join(hwm_dec_2022, bankfull), aes(x = station_1d, y = w_s_elev_ft ), color = "black", linetype="dotted") 

ggsave(file="out/lp_1d_thalweg.svg", plot=last_plot(), width=9, height=6)

```
Table comparing the RAS 1D result
```{r message=FALSE, warning=FALSE}
ras_1d %>% 
  bind_rows(ras_1d_afterstorm, .id="geometry") %>%
  select(geometry, cross_section, profile, w_s_elev_ft, min_ch_el_ft) %>%
  pivot_wider(names_from = geometry, values_from = c(min_ch_el_ft, w_s_elev_ft)) %>%
  unnest() %>%
  drop_na(cross_section) %>%
  mutate(w_s_elev_ft_delta = w_s_elev_ft_2 - w_s_elev_ft_1,
         min_ch_el_ft_delta = min_ch_el_ft_2 - min_ch_el_ft_1
         ) %>%
  filter(profile %in% c("Q2 BKF", "Q100 AC")) #%>% write_csv("out/ras_1d_comparison.csv")

```

Plots of RAS 1D results for each cross section individually

```{r message=FALSE, warning=FALSE}
geom_xs_wse <- function(data, water_elev, profile) {
  wse_data <- data %>% 
    arrange(sta) %>%
    mutate(wse = case_when(water_elev > gse ~ water_elev)) 
  return(geom_line(data=wse_data, aes(x = sta, y = wse, color = profile)))
}

gen_plot <- function(xs_df, xs_name) {
plot_xs(xs_df, 0) + 
  geom_xs_wse(xs_df, profile = "Q100 AC", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "Q100 AC") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "Q100 BKF", water_elev = ras_1d %>% filter(cross_section ==  xs_name & profile == "Q100 BKF") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "14 Jan 2023", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "14 Jan 2023") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "31 Dec 2022", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "31 Dec 2022") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "Q25 BKF", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "Q25 BKF") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "Q5 Channel", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "Q5 Channel") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "Q2 BKF", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "Q2 BKF") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "13 Feb 2019", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "13 Feb 2019") %>% pull(w_s_elev_ft)) + 
  geom_xs_wse(xs_df, profile = "11 Nov 2022", water_elev = ras_1d %>% filter(cross_section == xs_name & profile == "11 Nov 2022") %>% pull(w_s_elev_ft)) + 
    scale_color_manual(values = profile_colors, name="") + 
    theme(legend.position = "none") + 
    ylab("Elevation (ft NGVD29)") + 
    xlab("Station (ft from left bank)") + 
    ggtitle(paste0("Cross Section ",xs_name)) 
    
}
gen_plot(xs_b, "B") + coord_fixed(ratio=5, ylim=c(350,375), xlim=c(0,250), expand=FALSE)
ggsave(file=paste0("out/xs_b_1d.svg"), plot=last_plot(), width=9, height=4) 
gen_plot(xs_d, "D") + coord_fixed(ratio=5, ylim=c(345,370), xlim=c(0,250), expand=FALSE)
ggsave(file=paste0("out/xs_d_1d.svg"), plot=last_plot(), width=9, height=4)
gen_plot(xs_e, "E") + coord_fixed(ratio=5, ylim=c(340,365), xlim=c(-50,200), expand=FALSE)
ggsave(file=paste0("out/xs_e_1d.svg"), plot=last_plot(), width=9, height=4)
gen_plot(xs_f, "F") + coord_fixed(ratio=5, ylim=c(340,365), xlim=c(0,250), expand=FALSE)
ggsave(file=paste0("out/xs_f_1d.svg"), plot=last_plot(), width=9, height=4)
gen_plot(xs_g, "G") + coord_fixed(ratio=5, ylim=c(335,360), xlim=c(0,250), expand=FALSE)
ggsave(file=paste0("out/xs_g_1d.svg"), plot=last_plot(), width=9, height=4)
gen_plot(xs_h, "H") + coord_fixed(ratio=5, ylim=c(335,360), xlim=c(0,250), expand=FALSE)
ggsave(file=paste0("out/xs_h_1d.svg"), plot=last_plot(), width=9, height=4)

```
Model validation against high water marks

```{r message=FALSE, warning=FALSE}
# model validation
ras_1d %>% 
  left_join(hwm_geometries) %>% 
  filter(profile != "5 Jan 1982") %>%
  drop_na(water_surface_elevation) %>% 
  select(cross_section, series, profile, water_surface_elevation, w_s_elev_ft) %>%
  mutate(error = w_s_elev_ft - water_surface_elevation) %>%
  select(cross_section, series, profile, error) %>%
  arrange(cross_section) %>%
  pivot_wider(names_from = cross_section, values_from = error) %>%
  arrange(series, profile) #%>% write_csv("ignore/validation.csv")
``` 


# RAS 2D results for 31 Dec 2022 and Q100

Import RAS 2D results

```{r message=FALSE, warning=FALSE}

# RAS 2D results
ras_2d_gse <- read_csv("data/hec_ras_2d_profile_terrain.csv") %>%
  janitor::clean_names()

ras_2d_wse <- read_csv("data/hec_ras_2d_profile_wse_max.csv") %>%
  janitor::clean_names()

ras_2d_wse_q100 <- read_csv("data/hec_ras_2d_profile_wse_max_q100.csv") %>%
  janitor::clean_names()

ras_2d_gse_bda <- read_csv("data/hec_ras_2d_profile_terrain_bda.csv") %>%
  janitor::clean_names()

ras_2d_wse_bda <- read_csv("data/hec_ras_2d_profile_wse_max_bda.csv") %>%
  janitor::clean_names()

ras_2d_wse_q100_bda <- read_csv("data/hec_ras_2d_profile_wse_max_q100_bda.csv") %>%
  janitor::clean_names()

# Upper bank elevations calculated in ArcGIS
bank_elev <- inner_join(
    read_csv("data/bank_elev_lb.csv") %>%
      janitor::clean_names() %>% 
      rename(station_2d = objectid, bank_elev_left = elev) %>%
      select(station_2d, bank_elev_left),
    read_csv("data/bank_elev_rb.csv") %>%
      janitor::clean_names() %>% 
      rename(station_2d = objectid, bank_elev_right = elev) %>%
      select(station_2d, bank_elev_right)
  ) %>%
  mutate(bank_elev_min = case_when(
    bank_elev_left < bank_elev_right ~ bank_elev_left, 
    TRUE ~bank_elev_right)) %>%
  mutate(bank_elev_rolling_med = zoo::rollapply(bank_elev_min, 100, median, partial=T))
```

Plot RAS 2D long profile with high water marks

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = ras_2d_gse, aes(x = station_ft, y = terrain_profile_ft, color = "Terrain")) + 
  tidyquant::geom_ma(data = bank_elev, aes(x = station_2d, y = bank_elev_min, color = "Terrain"), 
                     ma_fun = EMA, n = 100, linetype = "dotted") +
  geom_point(data = hwm_dec_2022, aes(x = station_2d, y = hwm_elevation, color = "31 Dec 2022"), shape = 9, size = 4) + 
  geom_line(data = ras_2d_wse, aes(x = station_ft, y = wse_max_feet, color = "31 Dec 2022")) +
  geom_line(data = ras_2d_wse_q100, aes(x = station_ft, y = wse_max_feet, color = "Q100 AC")) +
#  geom_line(data = ras_2d_gse_bda, aes(x = station_ft, y = terrain_profile_ft, color = "Terrain Profile"), #linetype = "dashed") + 
#  geom_line(data = ras_2d_wse_bda, aes(x = station_ft, y = wse_max_feet, color = "Water Surface (RAS)"), #linetype = "dashed") +
  scale_color_manual(values = c("Terrain" = "black", 
                                "31 Dec 2022" = "#499894",
                                "Q100 AC" = "#ffae34")) + 
  theme(legend.position = "none") + 
  #ggtitle("2D model result") + 
  xlab("Station (ft from upstream boundary)") + 
  ylab("Elevation (ft NGVD29)") +
  theme(legend.position = "none") + #theme(legend.position = "right", legend.title=element_blank()) + 
  geom_text(data = hwm_dec_2022, aes(x = station_2d, y = thalweg_elevation - 1, color = "Terrain", label = cross_section )) 

ggsave(file="out/lp_2d_wse.svg", plot=last_plot(), width=9, height=6)
  
```

Now comparing original vs BDA modified terrain results

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = ras_2d_gse, aes(x = station_ft, y = terrain_profile_ft, color = "Terrain")) + 
  tidyquant::geom_ma(data = bank_elev, aes(x = station_2d, y = bank_elev_min, color = "Terrain"), 
                     ma_fun = EMA, n = 100, linetype = "dotted") +
  #geom_point(data = hwm_dec_2022, aes(x = station_2d, y = hwm_elevation, color = "31 Dec 2022"), shape = 9, size = 4) + 
  geom_line(data = ras_2d_wse, aes(x = station_ft, y = wse_max_feet, color = "31 Dec 2022")) +
  geom_line(data = ras_2d_wse_q100, aes(x = station_ft, y = wse_max_feet, color = "Q100 AC")) +
  geom_line(data = ras_2d_gse_bda, aes(x = station_ft, y = terrain_profile_ft, color = "Terrain Profile"), linetype = "dashed") + 
  geom_line(data = ras_2d_wse_bda, aes(x = station_ft, y = wse_max_feet, color = "31 Dec 2022"), linetype = "dashed") +
  geom_line(data = ras_2d_wse_q100_bda, aes(x = station_ft, y = wse_max_feet, color = "Q100 AC"), linetype = "dashed") +
    scale_color_manual(values = c("Terrain" = "black", 
                                "31 Dec 2022" = "#499894",
                                "Q100 AC" = "#ffae34")) + 
  theme(legend.position = "none") + 
  #ggtitle("2D model result") + 
  xlab("Station (ft from upstream boundary)") + 
  ylab("Elevation (ft NGVD29)") +
  theme(legend.position = "none") + #theme(legend.position = "right", legend.title=element_blank()) + 
  geom_text(data = hwm_dec_2022, aes(x = station_2d, y = thalweg_elevation - 1, color = "Terrain", label = cross_section )) 

ggsave(file="out/lp_2d_wse_bda.svg", plot=last_plot(), width=9, height=6)


```

Model validation 

```{r message=FALSE, warning=FALSE}
ras_2d_wse %>% 
  # join to the cross section table
  mutate(station_2d = as.integer(station_ft)) %>%
  group_by(station_2d) %>% 
  summarize(wse_2d = mean(wse_max_feet)) %>%
  inner_join(cross_sections) %>%
  inner_join(hwm_geometries %>% filter(peak_flow_date == ymd("2022-12-31"))) %>%
  select(cross_section, hwm_elevation, wse_2d) %>%
  mutate(error = wse_2d - hwm_elevation) %>%
  pivot_longer(cols = c(hwm_elevation, wse_2d, error)) %>%
  pivot_wider(names_from = cross_section, values_from = value) #%>% write_csv("ignore/validation2d.csv")
```

At-a-station hydraulics based on surveyed cross section geometries with RAS 2D water surface estimates


```{r message=FALSE, warning=FALSE}
# summarize at-a-station hydraulics for 2D results

ras_2d_cleaned <- ras_2d_wse %>% 
  # join to the cross section table
  mutate(station_2d = as.integer(station_ft)) %>%
  group_by(station_2d) %>% 
  summarize(wse_2d = mean(wse_max_feet)) %>% 
  mutate(slope = abs((lag(wse_2d, n = 1) - wse_2d) / 
                    (lag(station_2d, n = 1) - station_2d))) 

ras_2d_hydraulics <- ras_2d_cleaned %>%
  inner_join(cross_sections) %>%
  left_join(xs_dfs, by = join_by(cross_section == xs_id)) %>%
  # calculate hydraulic geometries
  mutate(result = map2(xs_df, wse_2d, calc_xs)) %>% 
  unnest_wider(result) %>% 
  # compare against surveyed HWMs
  left_join(hwm_dec_2022 %>% select(cross_section, hwm_elevation)) %>%
  mutate(hwm_difference = wse_2d - hwm_elevation)

ras_2d_hydraulics %>%
    rename(cross_sectional_area_ft2 = cross_sectional_area,
         wetted_perimeter_ft = wetted_perimeter) %>%
  mutate(discharge_cfs = peak_flow_dec_2022,
         velocity_ft_s = discharge_cfs / cross_sectional_area_ft2,
         hydraulic_radius_ft = cross_sectional_area_ft2 / wetted_perimeter_ft,
         mannings_n = 1.486 * cross_sectional_area_ft2 * hydraulic_radius_ft^(2/3) 
                      * slope^(1/2) * discharge_cfs^(-1),
         # metric conversions
         velocity_m_s = velocity_ft_s / 0.3048,
         hydraulic_radius_m = hydraulic_radius_ft / 0.3048,
         cross_sectional_area_m2 = cross_sectional_area_ft2 / 0.3048,
         # bed mobilization
         critical_shields_number = 0.15 * slope^(1/4),
         grain_size_mobilized_mm = 10 * rho_cgs * hydraulic_radius_m * slope / 
                         (critical_shields_number * (rho_s_cgs - rho_cgs)),
         grain_size_mobilized_phi = -log2(grain_size_mobilized_mm),
         # suspended transport
         shear_velocity_cm_s = sqrt(g_cgs * (hydraulic_radius_m*100) * slope),
         settling_velocity_ndim = rho_cgs * shear_velocity_cm_s^3 / 
                         ((rho_s_cgs - rho_cgs) * g_cgs * nu_cgs),
         grain_size_suspended_ndim = sqrt(5832 * settling_velocity_ndim),
         grain_size_suspended_mm = 10 * grain_size_suspended_ndim * rho_cgs * nu_cgs^2 /
                         ((rho_s_cgs - rho_cgs) * g_cgs)^(1/3),
         grain_size_suspended_phi = -log2(grain_size_suspended_mm)
  )

```

# Cross sections after Winter 2022-23 storms

Import cross section data after storm

```{r message=FALSE, warning=FALSE}

xs_b_afterstorm <- read_csv("data/xs_geom/xs_b_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_d_afterstorm <- read_csv("data/xs_geom/xs_d_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_e_afterstorm <- read_csv("data/xs_geom/xs_e_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_f_afterstorm <- read_csv("data/xs_geom/xs_f_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_g_afterstorm <- read_csv("data/xs_geom/xs_g_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

xs_h_afterstorm <- read_csv("data/xs_geom/xs_h_afterstorm.csv") %>%
  prep_xs(sta = station_left_bank_ft, elev = elevation_ngvd29_ft, delta_x = 0.1) 

tribble(~cross_section, ~xs_area_before_storm, ~xs_area_after_storm,
 "B",
calc_xs(xs_b, filter(hwm_dec_2022, cross_section=="B") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_b_afterstorm, filter(hwm_dec_2022, cross_section=="B") %>% pull(water_surface_elevation))$cross_sectional_area,
"D",
calc_xs(xs_d, filter(hwm_dec_2022, cross_section=="D") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_d_afterstorm, filter(hwm_dec_2022, cross_section=="D") %>% pull(water_surface_elevation))$cross_sectional_area,
"E",
calc_xs(xs_e, filter(hwm_dec_2022, cross_section=="E") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_e_afterstorm, filter(hwm_dec_2022, cross_section=="E") %>% pull(water_surface_elevation))$cross_sectional_area,
"F",
calc_xs(xs_f, filter(hwm_dec_2022, cross_section=="F") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_f_afterstorm, filter(hwm_dec_2022, cross_section=="F") %>% pull(water_surface_elevation))$cross_sectional_area,
"G",
calc_xs(xs_g, filter(hwm_dec_2022, cross_section=="G") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_g_afterstorm, filter(hwm_dec_2022, cross_section=="G") %>% pull(water_surface_elevation))$cross_sectional_area,
"H",
calc_xs(xs_h, filter(hwm_dec_2022, cross_section=="H") %>% pull(water_surface_elevation))$cross_sectional_area,
calc_xs(xs_h_afterstorm, filter(hwm_dec_2022, cross_section=="H") %>% pull(water_surface_elevation))$cross_sectional_area
) %>% mutate(xs_area_delta = xs_area_after_storm - xs_area_before_storm)
```

Plot depth/WSE versus discharge rating curves for before versus after storm
```{r message=FALSE, warning=FALSE}
ggplot() +
  geom_line(data = drop_na(ras_1d, cross_section), 
            aes(y = (w_s_elev_ft), x = q_total_cfs, linetype="Spring/Summer 2022")) + 
  geom_line(data = drop_na(ras_1d_afterstorm, cross_section), 
            aes(y = (w_s_elev_ft), x = q_total_cfs, linetype="Winter 2022-23"))  + 
  ylab("Water Surface Elevation (ft)") + xlab("Discharge (cfs)") + theme(legend.title=element_blank(), legend.position="top") + facet_wrap(~cross_section)

ggsave(file="out/ras_1d_ratingcurves_beforeafter.svg", plot=last_plot(), width=9, height=6)
```

